\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2024-02-09 RefereeResponse/response/RefereeResponse.tex

\date{\today}

\maketitle
Dear Editor,

\

We would like to begin by sincerely thanking the referee for their careful reading of the manuscript.
We believe their thoughtful comments and suggestions have produced a better paper.
The referee suggested more background for the two-loop results, and noted that a comparison with $\mathcal{N}=4$ super-Yang Mills (sYM) was absent. We agree that adding this context provides a necessary framing of the main technical results.
To this end, below is a summary of the modifications to the manuscript:
\begin{itemize}
\item A small comment has been added to the introduction at the top of page 2 reminding readers that color-dual representations of sYM have been found to high loop order.
\item At the bottom of page 2, when stating that there is no local 4pt 2-loop numerator for pure YM, we remind readers that such a numerator exists for sYM.
\item At the end of the introduction at the top of page 3 we clarify that the bowtie cut failure is specific to pure YM.
We go on to speculate that this is because sYM and NLSM are simpler theories, at least as far as color-kinematics is concerned.
\item In addition to the comments in the introduction and section 4, we have included one additional modification.
Section 2.3 deals with ansatz locality so this seemed like a natural place to mention that the sYM ans\"atze incorporate locality as a key simplifying assumption.
At the bottom of page 7 and the top of page 8, we emphasize that the sYM numerators through four loops are local.
We comment in footnote 4 that the full set of local numerators has never been fully explored for sYM at 4pt 5-loop.
\item The first paragraph in section 4 closes by putting our results in the context of sYM, specifically by reminding readers of the state of the art in that theory. Another such reminder is included in footnote 10 at the beginning of section 4.3. 
\item After stating the bowtie failure condition for pure YM in (4.9) we have added a paragraph explaining why NLSM and sYM are more privileged with regards to color-kinematics duality and the bowtie cut. 
\end{itemize}

We would like to thank the referee again for their carefully considered feedback.
Now that the referee's comments have been addressed and the readability of the manuscript has been improved, we believe that the paper would be an excellent fit for publication in JHEP.

### 2024-02-09 RefereeResponse/paper/2loopNLSM.tex
Small referee edits

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric
theories.
However, in sharp contrast to both sYM \cite{BRY} and NLSM, we find that Yang-Mills does %Referee edit.  Am I citing the right paper?
\emph{not} permit a globally color-dual two-loop representation, even when
considering the most general polynomial ansatz of Lorentz covariant
kinematics.\footnote{Globally color-dual local numerators for two-loop
  pure YM have been studied before but only in the context of a
  limited ansatz subject to certain loop power counting constraints
  \cite{Bern:2015ooa}.}
  Concretely, the failure point can be
pinpointed to a conflict between the ``bowtie'' cut and the
following Jacobi triple:
\begin{equation}
\label{eq:IntroBowtieFailure}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
This failure is specific to Yang-Mills and does not occur for its maximally supersymmetric counterpart or for NLSM, indicating that pure Yang-Mills is a more complicated theory in some sense. %Referee edit
We comment on the implications of \cref{eq:IntroBowtieFailure} in \cref{conclusions}
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual integrands at generic loop order.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states. Locality is physically motivated
since Feynman rules produce polynomial numerators.  Indeed, the most
natural way to guarantee locality and avoid spurious poles is to
require the off-shell numerators to be polynomial functions of
kinematics.  However, insisting on a local numerator is partly a
matter of convenience as the ansatz for a rational function would
quickly grow out of control without a guiding principle for what terms
to include.  %Referee edit
Historically, locality has also proved a highly successful tool for
simplifying kinematic
numerators. %Alex, can you add more relevant references to the next two sentences?
For example, the four-point function for $\mathcal{N}=4$ super-Yang
Mills admits local color-dual numerators through four loops
\cite{Bern:2012uf}.\footnote{At five loops, the most general local
  ansatz has not been explored, but the double copy has been performed
  at five loops using alternative methods \cite{GeneralizedDoubleCopy,
    GeneralizedDoubleCopyFiveLoops}.}  While locality is
well motivated from a physical and complexity standpoint, the
literature has featured numerators that violate locality, modify naive
power counting through the inclusion of higher spin modes, and break
manifest Lorentz invariance \cite{Square, WeinzierlBCJLagrangian,
  Mogull:2015adi, FivePointN4BCJ, Johansson:2017bfl,
  Ben-Shahar:2022ixa, Cheung:2016prv, Cheung:2021zvb}.  These
developments suggest that relaxing locality is a viable alternative
when the condition proves too stringent for an ansatz.  However,
simply requiring on-shell locality may suffice.  This alternative will
be discussed in \cref{nonLocalScattering}.

@@

\section{Two-loop four-point bootstrap}
\label{2loopBoot}
As we have just seen, it is possible to coerce tree numerators into
one-loop numerators at any multiplicity for pions and related
theories, but that these methods cannot be reapplied to generate
higher-loop numerators.  For pions in particular, the two-loop no-go
statement is only for \emph{one particular representation of the
  theory}, so it does not completely preclude the existence of a
two-loop color-dual integrand.  We thus turn to the color-dual
bootstrap method described in \cref{sec:bootstrap} to construct a
color-dual representation of two-loop four-point NLSM.  Because of the
similarity of the problem setup, we will also use the opportunity to
revisit the construction of a color-dual representation of pure YM,
extending the search space beyond what was covered in
Ref.~\cite{Bern:2015ooa} to include the most general local ansatz.
Even though this work is concerned with non-supersymmetric theories,
recall that \ansatze{} techniques have been very successful for
sYM.  %Referee edit
In particular, local color-dual representations have been found
through 4pt 4-loop \cite{Bern:2012uf}. %What paper should we cite?
While a local representation has not been found at five loops, the
double copy has still be performed
\cite{GeneralizedDoubleCopy,GeneralizedDoubleCopyFiveLoops}.

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of
Ref.~\cite{Bern:2015ooa}.\footnote{As mentioned at the beginning of
  this section, \emph{maximally supersymmetric} Yang-Mills is known to have a
  local color-dual representation at least through four loops
  \cite{Bern:2012uf}.  A local but restricted ansatz fails at five loops, but
  the most general local ansatz has not been explored
  \cite{GeneralizedDoubleCopyFiveLoops}.} %Referee edits.  What papers should we cite?
We follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box and penta-triangle as the Jacobi basis
graphs and building the most general ansatz for each of their
numerators compatible with the assumptions in \cref{sec:bootstrap}.
Because we are now considering pure Yang-Mills, each monomial in the
local ansatz must consist of five Lorentz scalar dot products instead
of the six for NLSM, and each monomial must be linear in each of the
four external gluon polarizations.  Thus, the numerator ansatz for
both diagrams will be built from terms of the form
\begin{equation}
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisDoubleBox} \right]
  \text{ and }
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]
  \in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  Without any power counting restrictions
imposed, both the double-box and the penta-triangle have an ansatz
with 10,010 terms each.  Imposing diagram automorphisms and
maximal-cut gauge invariance on the two basis diagrams, we reduce the
number of terms to 2,235 for the double-box, and 4,133 for the
penta-triangle.  The minimal set of spanning physical cuts is shown in
\cref{fig:ym-spanning}, but we choose to work within the framework of
the method of maximal cuts \cite{Bern:2007ct} in order to identify the
simplest cut that is in tension with the kinematic Jacobi relations.
Maximal cuts and symmetries are then imposed on the 14
non-pathological cubic diagrams shown above in \cref{fig:MaxCuts}.
Doing so, we are left with 596 total parameters in the ansatz.

@@

Concretely, in this work we have found that the power counting of NLSM offers sufficient generalized gauge freedom to manifest the duality globally, while maintaining compatibility with the ``bowtie'' cut. Whereas in the case of $\mathcal{N}=4$ sYM, by virtue of permitting a local construction consistent with \cref{chefsKiss}, the supersymmetric state sum\footnote{Due to the presence of fermionic modes that propagate off-shell in $\mathcal{N}=4$ sYM amplitudes, some of the pure vector contributions will be cancelled in the ``bow-tie'' cut, which leads to simpler cut expressions \cite{SuperSum}.} must be sufficiently simple, in such a way that it absorbs cut contributions present in the pure gluon state sum that we have shown are inconsistent with a globally color-dual polynomial ansatz. As such, the result of \cref{chefsKiss} is due to the power counting of pure YM being too restrictive to account for the complexity posed by the pure gluon ``bow-tie'' cut.  %Referee edit

>>>>

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric
theories.
However, in sharp contrast to both sYM \cite{BRY} and NLSM, we find that Yang-Mills does %Referee edit
\emph{not} permit a globally color-dual two-loop representation, even when
considering the most general polynomial ansatz of Lorentz covariant
kinematics.\footnote{Globally color-dual local numerators for two-loop
  pure YM have been studied before but only in the context of a
  limited ansatz subject to certain loop power counting constraints
  \cite{Bern:2015ooa}.}
  Concretely, the failure point can be
pinpointed to a conflict between the ``bowtie'' cut and the
following Jacobi triple:
\begin{equation}
\label{eq:IntroBowtieFailure}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
This failure is specific to Yang-Mills and does not occur for its maximally supersymmetric counterpart or for NLSM, indicating that pure Yang-Mills is a more complicated theory in some sense. %Referee edit
We comment on the implications of \cref{eq:IntroBowtieFailure} in \cref{conclusions}
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual integrands at generic loop order.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states. Locality is physically motivated
since Feynman rules produce polynomial numerators.  Indeed, the most
natural way to guarantee locality and avoid spurious poles is to
require the off-shell numerators to be polynomial functions of
kinematics.  However, insisting on a local numerator is partly a
matter of convenience as the ansatz for a rational function would
quickly grow out of control without a guiding principle for what terms
to include.  %Referee edit
Historically, locality has also proved a highly successful tool for
simplifying kinematic
numerators.
For example, the four-point function for $\mathcal{N}=4$ super-Yang
Mills admits local color-dual numerators through four loops
\cite{Bern:2012uf}.\footnote{At five loops, the most general local
  ansatz has not been explored, but the double copy has been performed
  at five loops using alternative methods \cite{GeneralizedDoubleCopy,
    GeneralizedDoubleCopyFiveLoops}.}  While locality is
well motivated from a physical and complexity standpoint, the
literature has featured numerators that violate locality, modify naive
power counting through the inclusion of higher spin modes, and break
manifest Lorentz invariance \cite{Square, WeinzierlBCJLagrangian,
  Mogull:2015adi, FivePointN4BCJ, Johansson:2017bfl,
  Ben-Shahar:2022ixa, Cheung:2016prv, Cheung:2021zvb}.  These
developments suggest that relaxing locality is a viable alternative
when the condition proves too stringent for an ansatz.  However,
simply requiring on-shell locality may suffice.  This alternative will
be discussed in \cref{nonLocalScattering}.

@@

\section{Two-loop four-point bootstrap}
\label{2loopBoot}
As we have just seen, it is possible to coerce tree numerators into
one-loop numerators at any multiplicity for pions and related
theories, but that these methods cannot be reapplied to generate
higher-loop numerators.  For pions in particular, the two-loop no-go
statement is only for \emph{one particular representation of the
  theory}, so it does not completely preclude the existence of a
two-loop color-dual integrand.  We thus turn to the color-dual
bootstrap method described in \cref{sec:bootstrap} to construct a
color-dual representation of two-loop four-point NLSM.  Because of the
similarity of the problem setup, we will also use the opportunity to
revisit the construction of a color-dual representation of pure YM,
extending the search space beyond what was covered in
Ref.~\cite{Bern:2015ooa} to include the most general local ansatz.
Even though this work is concerned with non-supersymmetric theories,
recall that \ansatze{} techniques have been very successful for
sYM.  %Referee edit
In particular, local color-dual representations have been found
through 4pt 4-loop \cite{Bern:2012uf}.
While a local representation has not been found at five loops, the
double copy has still been performed
\cite{GeneralizedDoubleCopy,GeneralizedDoubleCopyFiveLoops}.

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of
Ref.~\cite{Bern:2015ooa}.\footnote{As mentioned at the beginning of
  this section, \emph{maximally supersymmetric} Yang-Mills is known to have a
  local color-dual representation at least through four loops
  \cite{Bern:2012uf}.  A local but restricted ansatz fails at five loops, but
  the most general local ansatz has not been explored
  \cite{GeneralizedDoubleCopyFiveLoops}.} %Referee edit
We follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box and penta-triangle as the Jacobi basis
graphs and building the most general ansatz for each of their
numerators compatible with the assumptions in \cref{sec:bootstrap}.
Because we are now considering pure Yang-Mills, each monomial in the
local ansatz must consist of five Lorentz scalar dot products instead
of the six for NLSM, and each monomial must be linear in each of the
four external gluon polarizations.  Thus, the numerator ansatz for
both diagrams will be built from terms of the form
\begin{equation}
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisDoubleBox} \right]
  \text{ and }
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]
  \in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  Without any power counting restrictions
imposed, both the double-box and the penta-triangle have an ansatz
with 10,010 terms each.  Imposing diagram automorphisms and
maximal-cut gauge invariance on the two basis diagrams, we reduce the
number of terms to 2,235 for the double-box, and 4,133 for the
penta-triangle.  The minimal set of spanning physical cuts is shown in
\cref{fig:ym-spanning}, but we choose to work within the framework of
the method of maximal cuts \cite{Bern:2007ct} in order to identify the
simplest cut that is in tension with the kinematic Jacobi relations.
Maximal cuts and symmetries are then imposed on the 14
non-pathological cubic diagrams shown above in \cref{fig:MaxCuts}.
Doing so, we are left with 596 total parameters in the ansatz.

@@

Concretely, in this work we have found that the power counting of NLSM offers sufficient generalized gauge freedom to manifest the duality globally, while maintaining compatibility with the ``bowtie'' cut. Whereas in the case of $\mathcal{N}=4$ sYM, by virtue of permitting a local construction consistent with \cref{chefsKiss}, the supersymmetric state sum\footnote{Due to the presence of fermionic modes that propagate off-shell in $\mathcal{N}=4$ sYM amplitudes, some of the pure vector contributions will be cancelled in the ``bowtie'' cut, which leads to simpler cut expressions \cite{SuperSum}.} must be sufficiently simple, in such a way that it absorbs cut contributions present in the pure gluon state sum that we have shown are inconsistent with a globally color-dual polynomial ansatz. As such, the result of \cref{chefsKiss} is due to the power counting of pure YM being too restrictive to account for the complexity posed by the pure gluon ``bowtie'' cut.  %Referee edit
